ML-based approach for accelerating global search algorithm for solving multicriteria problems\thanks{This study was supported by the Russian Science Foundation, project No 21-11-00204.}

### 2022-04-29 2022/LION/main.tex
Lion добавлен график изменения HV при разном числе итераций
Обновил график
include three new rerefences

\keywords{Multicriterial Problems \and Global Optimization \and Machine Learning \and Logistic Regression.}
\end{abstract}
%
%
%
\section{Introduction}
Machine learning (ML) methods are powerful tools that are applied in many areas of research.
In particular, the ML methods are successfully used for solving complicated problems of computational mathematics. One of such problem classes to which ML can be applied is the class of multicriterial optimization (MCO) models. 
In this case, ML is used, as a rule, in combination with metaheuristic algorithms \cite{Subraveti2019} that in the case of multiextremal criteria lose in efficiency to deterministic methods \cite{Kvasov2018,Sergeyev2018}.

@@

Computational testing was carried out on the``Lobachevsky'' supercomputer of Nizhny Novgorod State University in the environment of the Globalizer software system \cite{globalizerSystem}.

@@

In order to estimate efficiency of machine learning procedure in ML\_MGSA, this method tested in 2 variants: with ($\alpha>0$) and without ($\alpha=0$)
use of separating planes in the criteria space. 
Each multicriterial problem was reduced to 50 scalar problems (\ref{eq:4}) with different weight coefficients $\lambda$, distributed uniformly. Other algorithm's parameters were the same. The averaged results can be found in Table \ref{tab:2}.

@@

As shown in the table, embedding the machine learning into the global search algorithm results in a significant reduction of the number of trials (almost 3 times) with maintaining close values of indicators HV and DU.

>>>>

\keywords{Multicriterial Problems \and Global Optimization \and Machine Learning \and Logistic Regression.}
\end{abstract}
%
%
%
\section{Introduction}
Machine learning (ML) methods are powerful tools that are applied in many areas of research.
In particular, the ML methods are successfully used for solving complicated problems of computational mathematics. One of such problem classes to which ML can be applied is the class of multicriterial optimization (MCO) models. 
In this case, ML is used, as a rule, in combination with metaheuristic algorithms \cite{Ma2014,Subraveti2019,Talbi2015} that in the case of multiextremal criteria lose in efficiency to deterministic methods \cite{Kvasov2018,Sergeyev2018}.

@@

Computational testing was carried out on the ``Lobachevsky'' supercomputer of Nizhny Novgorod State University in the environment of the Globalizer software system \cite{globalizerSystem}.

@@

In order to estimate efficiency of machine learning procedure in ML\_MGSA, this method tested in 2 variants: with ($\alpha>0$) and without ($\alpha=0$)
use of separating planes in the criteria space. 
Each multicriterial problem was reduced to 50 scalar problems (\ref{eq:4}) with different weight coefficients $\lambda$, distributed uniformly. Other algorithm's parameters were the same. The averaged results can be found in Table~\ref{tab:2}. \textcolor[rgb]{1,0,0}{Fig.~\ref{fig:2} additionally shows the average value of HV index change depending of the number of trials.}

@@

\begin{figure}[ht]
\center
\includegraphics[width=0.7\textwidth]{fig2.png}
\caption{Efficiency of the multicriterial algorithms compared} \label{fig:2}
\end{figure}

\textcolor[rgb]{1,0,0}{The results show that embedding the machine learning into the global search algorithm leads to }a significant reduction of the number of trials (almost 3 times) with maintaining close values of indicators HV and DU.
